### 2023-01-16 experiments/superg/doc/paper/paper.tex

\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage[pdftex]{graphicx}
% declare the path(s) where your graphic files are
\graphicspath{{images/}}

% Packages that are sometimes useful
%\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
%\usepackage{float}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Reviving the Lost Art of Combinator Graph Reduction},
    pdfpagemode=FullScreen,
    }

\urlstyle{same}

\usepackage{amsmath}
\DeclareMathOperator{\strict}{strict}
\DeclareMathOperator{\lazy}{lazy}
\DeclareMathOperator{\lazyeta}{lazy_\eta}
\DeclareMathOperator{\linear}{linear}

\usepackage{listings}

\begin{document}

\title{Reviving the Lost Art of Combinator Graph Reduction}

\author{\IEEEauthorblockN{Daan de Graaf}
    \IEEEauthorblockA{\textit{Master's student Computer Engineering} \\
        \textit{TU Delft / Eindhoven University of Technology}\\
        Eindhoven, Netherlands \\
        d.j.a.degraaf@student.tudelft.nl / d.j.a.d.graaf@student.tue.nl}
}

\maketitle

\begin{abstract}
    % Rules of thumb
    % - 1 sentence background
    % - 1 sentence motive/problem
    % - 1 sentence objective
    % - 1 sentence approach/method
    % - 2 sentences support
    % - 1 sentence conclusion
    % - 1 sentence implication
    % - 1 sentence limitation
    This is the abstract.
\end{abstract}

\begin{IEEEkeywords}
    combinatory logic, graph reduction, lazy functional language, lazy evaluation
\end{IEEEkeywords}

\section{Introduction}
% Why did I research this? Give the motive, usually a knowledge gap or problem
Lazy evaluation of programs is a powerful idea.
It allows a program to operate on infinitely large data structures, and if part of the computation turns out to be unnecessary, it is more efficient than a strict evaluation.
In practice, delaying computation often introduces overhead, making languages based on lazy evaluation slower than their strict counterparts.
To compete with a strict implementation, a lazy runtime must maximize laziness while minimizing overhead.
It is a difficult balance, and over the years many different evaluator designs, or \textit{abstract machines}, have been proposed~\cite{turner_new_1979,kieburtz_g-machine_1985,fairbairn_tim_1987,burn_spineless_1988,koopman_fresh_1989,jones_implementing_nodate}.

One of the early approaches to implementing lazy functional languages is combinator graph reduction~\cite{turner_new_1979}.
It was the basis for David Turner's Miranda language~\cite{turner_miranda_1985}, the popular choice for pure functional languages and one of the few to be commercially supported at the time.
Eventually, the interest in Miranda and combinator graph reduction with it waned.
Part of the reason may be that Miranda was until very recently a closed-source, licensed software (its source code was released only in 2020~\cite{noauthor_open_2021}), leading to the creation of the open-source Haskell language~\cite{hudak_history_2007}.
But combinator reduction also deserves part of the blame.
% Firstly, owing to the simplicity of individual combinators, expressions are reduced in many small steps, so the per-step overhead greatly impacts runtime performance.
% Every reduction step modifies the stack and usually also the graph, resulting in high memory traffic.
One of the major issues is that the translation to combinators may result in much larger programs.
For a lambda calculus expression of size $N$, the traditional bracket abstraction algorithm produces an output of $O(N^2)$ in the worst case, where other approaches such as supercombinators are linear\cite{spj_impl}.
Larger programs run slower, and along with other disadvantages, this led to combinator graph reduction being all but abandoned.

% What did I research?
% - Research problem/question
%   * Definitions
%   * Choices
%   * Presuppositions
% - Why specifically this research question?
%   * Theoretical or societal relevance
Recently, however, a new algorithm for the translation of lambda calculus expressions into combinatory logic has been proposed that is linear in the size of the input\cite{kiselyov_lambda_2018}.
Kiselyov's translation algorithm is based on semantics rather than syntax and this deeper understanding of the program has been shown to produce more compact translations.
While a clear improvement over previous translation strategies, it remains to be seen how big the impact is on real-world performance.
Is it enough to bridge the performance gap with newer reduction engines based on different paradigms?
This leads us to our first research question:

\textbf{RQ1:} Does Kiselyov's semantic translation make combinator graph reduction competitive with contemporary lazy functional evaluators?

% How did I research this?
% - Methods
% - Sub-questions
We use GHC as the reference for a contemporary evaluator, and then measure the execution time of a set of benchmark programs.
We also need a fast reduction engine that can run programs compiled to combinator expressions.
Alas, there is no performant reduction engine tailored to executing such programs.
The now open-sourced Miranda is a good starting point, but it is showing its age.
The 2020 port compiles cleanly on a recent GCC, but enabling optimizations is known to break the garbage collector.
Miranda's architecture is based on cells with a tag, head and tail as described in\cite{turner_new_1979}, but later work has shown that a tagless representation with direct pointers to code leads to better performance\cite{koopman_fresh_1989}.
In absence of an up-to-date reference on building a practical combinator reduction engine, we formulate the second research question:

\textbf{RQ2:} How is combinator graph reduction implemented efficiently on conventional hardware?

We evaluate the state of the art in combinator graph reduction and build a fast implementation for a language that is simple but powerful enough to express our benchmark programs.
Using this engine, we can then answer \textbf{RQ1}.

% How did I organize the report?
% - Structure Overview
Concretely, our contributions are the following:
\begin{itemize}
    \item Kiselyov develops not one but two algorithms. The first does not have linear complexity, but tends to produce very compact code.
          The second is guaranteed to be of linear complexity, at the cost of being less compact in certain cases.
          Kiselyov's description is somewhat theoretical, so in section~\ref{sec:kiselyov} we describe both algorithms in a way that is straightforward to implement.
    \item We implement a fast combinator graph reduction engine based on the ideas in Miranda~\cite{turner_miranda_1985} and later improvements~\cite{koopman_architecture_1992}.
          We elaborate on its design in sections~\ref{sec:miranda} and~\ref{sec:tigre}.
    \item We evaluate the performance of Miranda, Haskell and our engine on a set of benchmark programs in section~\ref{sec:eval}.
\end{itemize}

For readers unfamiliar with combinatory logic and/or graph reduction, we provide a short introduction to these topics in section~\ref{sec:prelim}.

\section{Preliminaries}
\label{sec:prelim}
\textbf{TODO: write section. Currently contains only a few small fragments}
\subsection{Combinatory Logic}
Starting from a lambda calculus expression, we first translate it into an expression consisting exclusively of the application of a small set of predefined functions, or \textit{combinators}.
It turns out that any lambda calculus expression can be expressed using just three combinators:
\begin{itemize}
    \item $S \ f \ g \ x = f \ x \ (g \ x)$
    \item $K \ x \ y = x$
    \item $I \ x = x$\footnote{It turns out that $SKK = I$, so even this combinator is not strictly necessary, but all reasonable implementations do include $I$, and it simplifies the example.}
\end{itemize}

For example, $\lambda x.x+1$ can be converted into the combinator expression:

\[
    S \ (S \  (K \  +) \  I) \ (K \ 1)
\]

We invite the reader to verify this by applying $x$ to this expression and reducing it to $x+1$ with the usual $\beta$-reduction rule.
While the generated expression is not as easy to read, the simplicity of its components is appealing: to execute a program compiled to combinators, an evaluator only needs to support four operations: the built-in $S$, $K$ and $I$ combinators, and function application\footnote{And any builtins of the language, like the plus operator in the example.}.
Better still, conversion to combinators removes all bound variables.
The runtime need not concern itself with passing arguments, this is all handled by the combinators.
The compiled program is executed using graph reduction, a technique that is also common in other reduction engines\cite{kieburtz_g-machine_1985,fairbairn_tim_1987}.

\subsection{Graph Reduction}
% In the reduction graph, vertices are combinators, constants or function applications, and edges connect the function and argument for application (combinators and constants have no outgoing edges, they are leaf nodes).

\section{Kiselyov's algorithms}
\label{sec:kiselyov}
Kiselyov's paper~\cite{kiselyov_lambda_2018} describes three iterations of his non-linear semantic translation, and finally the linear algorithm.
They are presented as a type system accompanied with operational semantics, with the exception of the $\eta$-optimization given as an OCaml code patch.
OCaml's type system can represent these algorithms very concisely in tagless-final style, but how should we represent them in a more conventional style, possibly in a language that does not have such a strong type system?

\subsection{Strict algorithm}
This is based on the rules in figure 6 of Kiselyov's paper.
We call it \textit{strict} it does not have lazy weakening: most notably it lacks the K-optimization required for full laziness.

The most naive implementation is to perform the type checking, then use the built context to apply the elimination rules and the translated program.
As Kiselyov already notes in his treatment of the linear algorithm further on in the paper, the translation does not actually care what elements are in the context, just that we keep track of the size of the context.
This lets us express compilation as a recursive function:

\begin{equation*}
    \begin{array}{l l l l}
        \strict & z         & = & 1 \models I                                              \\
        \strict & s \ e     & = & n+1 \models (0 \models K) \coprod (n \models c)          \\
                &           &   & \text{where } (n \models c) := \strict e                 \\
        \\
        \strict & \lambda e & = & 0 \models K \; c                                         \\
                &           &   & \text{where } (0 \models c) := \strict e                 \\
        \strict & \lambda e & = & (n-1) \models K \; c                                     \\
                &           &   & \text{where } (n \models c) := \strict e                 \\
        \\
        \strict & e_1 \ e_2 & = & (n \models ((n_1 \models c_1) \coprod (n_2 \models c_2)) \\
                &           &   & \text{where } (n_1 \models c_1) := \strict e_1           \\
                &           &   & \text{where } (n_2 \models c_2) := \strict e_2           \\
                &           &   & \text{where } n := \max n_1 \, n_2                       \\
    \end{array}
\end{equation*}

There are two cases for $\lambda e$, and this is precisely where the number of elements in the context matters:
if the context is empty, the first rule applies, otherwise, we use the second one.
The semantic function may be defined as:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        (0 \models c_1)   & \coprod & (0 \models c_2)   & = & c_1 \ c_2                                         \\
        \\
        (0 \models c_1)   & \coprod & (n_2 \models c_2) & = & (0 \models B c_1) \coprod (n_2 - 1 \models c_2)   \\
        \\
        (n_1 \models c_1) & \coprod & (0 \models c_2)   & = & (0 \models C C c_2) \coprod (n_1 - 1 \models c_1) \\
        \\
        (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & (n_1 - 1 \models (s \coprod l)) \coprod r)        \\
                          &         &                   &   & \text{where } s := (0 \models S )                 \\
                          &         &                   &   & \text{where } l := (n_1-1 \models c_1 )           \\
                          &         &                   &   & \text{where } r := (n_2-1 \models c_2 )           \\
    \end{array}
\end{equation*}

These simplified functions can be implemented directly and efficiently in a conventional programming language.

\subsection{Lazy weakening}
Adding lazy weakening requires that we track which items in the context are ignored.
We change our context representation from an integer to a list of boolean values.
In the new compilation functions below, we use a shorthand $t$ for the value \textit{true} and $f$ for \textit{false}.
We always match the last element of lists and represent the remainder of the list with $\Gamma$.

\begin{equation*}
    \begin{array}{l l l l}
        \lazy & z         & = & t \models I                                                             \\
        \lazy & s \ e     & = & \Gamma,f \models c                                                      \\
              &           &   & \text{where } (\Gamma \models c) := \lazy e                             \\
        \\
        \lazy & \lambda e & = & \emptyset \models K \; c                                                \\
              &           &   & \text{where } (\emptyset \models c) := \lazy e                          \\
        \lazy & \lambda e & = & (\emptyset \models K) \coprod (\Gamma \models c)                        \\
              &           &   & \text{where } (\Gamma,f \models c) := \lazy e                           \\
        \lazy & \lambda e & = & \Gamma \models K \; c                                                   \\
              &           &   & \text{where } (\Gamma,t \models c) := \lazy e                           \\
        \\
        \lazy & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)) \\
              &           &   & \text{where } (\Gamma_1 \models c_1) := \lazy e_1                       \\
              &           &   & \text{where } (\Gamma_2 \models c_2) := \lazy e_2                       \\
              &           &   & \text{where } \Gamma := \Gamma_1 \sqcup \Gamma_2                        \\
    \end{array}
\end{equation*}

We have introduced yet another case for $\lambda e$, this time to distinguish between $t$ or $f$ as the last element in the context of the inner expression.
There is also the new function $\sqcup$, which we need to implement:

\begin{equation*}
    \begin{array}{l l l l l}
        \Gamma_1    & \sqcup & \emptyset   & = & \Gamma_1                     \\
        \emptyset   & \sqcup & \Gamma_2    & = & \Gamma_2                     \\
        \Gamma_1,t  & \sqcup & \Gamma_2,\_ & = & (\Gamma_1 \sqcup \Gamma_2),t \\
        \Gamma_1,\_ & \sqcup & \Gamma_2,t  & = & (\Gamma_1 \sqcup \Gamma_2),t \\
        \Gamma_1,f  & \sqcup & \Gamma_2,f  & = & (\Gamma_1 \sqcup \Gamma_2),f \\
    \end{array}
\end{equation*}

In our implementation, we unroll the recursion into a while loop, and repeatedly pop elements from the two arrays until both are empty.
This builds the new context backwards, so as a final step the array is reversed.

Alternatively, an implementation can pad the shorter list (at the start!) with $f$ and zip them.
This is particularly convenient for languages that provide a zip function with a default value\footnote{such as \href{https://docs.python.org/3/library/itertools.html\#itertools.zip\_longest}{\texttt{ziplongest}} in Python.}.

Our semantic function now has to handle more cases, and thus is somewhat larger:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        (\emptyset \models c_1)  & \coprod & (\emptyset \models c_2)  & = & c_1 \ c_2                                                \\
        \\
        (\emptyset \models c_1)  & \coprod & (\Gamma_2,t \models c_2) & = & (\emptyset \models B c_1) \coprod (\Gamma_2 \models c_2) \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\emptyset \models C C c_2) \coprod (\Gamma \models c_1) \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (s \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } s := (\emptyset \models S )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        % The new additions
        (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)    \\
        \\
        (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (b \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } b := (\emptyset \models B )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models (c \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } c := (\emptyset \models C )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        (\emptyset \models c_1)  & \coprod & (\Gamma_2,f \models c_2) & = & (\emptyset \models c_1) \coprod (\Gamma_2 \models c_2)   \\
        \\
        (\Gamma_1,f \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\Gamma_1 \models c_1) \coprod (\emptyset \models c_2)   \\
    \end{array}
\end{equation*}

The first 4 cases are direct translations from the previous definitions, the others are new additions to handle ignored elements.
The last 2 cases are not included in the original paper, but they are necessary to make the pattern matching exhaustive, and are present in the reference OCaml implementation.

\subsection{Eta optimization}
Kiselyov covers eta optimization as a change to his OCaml code rather than typing rules.
Our recursive functions are derived from the typing rules, so these changes are not as straightforward to incorporate.
%Previously Kiselyov's representation consisted of elements $C$, $N$ and $W$, representing closed expressions, a used context element, and an ignored context element respectively.
The introduction of the $V$ element means that a context may now be associated with either a closed expression or the top variable (we reuse $V$ as a marker for this).
Our implementation uses a tagged union to define a type that can be either an expression or the top variable.
For the compilation function, we only need to change the definition of $\lazyeta \ z$ and add an extra case for $\lazyeta \ \lambda e$:

\begin{equation*}
    \begin{array}{l l l l}
        \lazyeta & z         & = & \emptyset \models V                                                          \\
        \lazyeta & s \ e     & = & \Gamma,f \models c                                                           \\
                 &           &   & \text{where } (\Gamma \models c) := \lazyeta e                               \\
        \\
        \lazyeta & \lambda e & = & \emptyset \models I                                                          \\
                 &           &   & \text{where } (\emptyset \models V) := \lazyeta e                            \\
        \lazyeta & \lambda e & = & \emptyset \models K \; c                                                     \\
                 &           &   & \text{where } (\emptyset \models c) := \lazyeta e                            \\
        \lazyeta & \lambda e & = & (\emptyset \models K) \coprod_\eta (\Gamma \models c)                        \\
                 &           &   & \text{where } (\Gamma,f \models c) := \lazyeta e                             \\
        \lazyeta & \lambda e & = & \Gamma \models K \; c                                                        \\
                 &           &   & \text{where } (\Gamma,t \models c) := \lazyeta e                             \\
        \\
        \lazyeta & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod_\eta (\Gamma_2 \models c_2)) \\
                 &           &   & \text{where } (\Gamma_1 \models c_1) := \lazyeta e_1                         \\
                 &           &   & \text{where } (\Gamma_2 \models c_2) := \lazyeta e_2                         \\
                 &           &   & \text{where } \Gamma := \Gamma_1 \sqcup_\eta \Gamma_2                        \\
    \end{array}
\end{equation*}

Our context merge function $\sqcup$ now also depends on the compiled expression, since we have a special case for $V$:

\begin{equation*}
    \begin{array}{l l l l l}
        (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\emptyset \models V)     & = & \Gamma_1,t                        \\
        (\emptyset \models V)     & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & \Gamma_2,t                        \\
        (\emptyset \models c_1)   & \sqcup_\eta & (\emptyset \models V)     & = & t                                 \\
        (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models c_2)   & = & t                                 \\
        (\emptyset \models V)     & \sqcup_\eta & (\emptyset \models V)     & = & \mathbf{impossible}               \\
        \\
        % Original ones
        (\Gamma_1 \models c_1)    & \sqcup_\eta & (\emptyset \models c_2)   & = & \Gamma_1                          \\
        (\emptyset \models c_1)   & \sqcup_\eta & (\Gamma_2 \models c_2)    & = & \Gamma_2                          \\
        (\Gamma_1,t \models c_1)  & \sqcup_\eta & (\Gamma_2,\_ \models c_2) & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
        (\Gamma_1,\_ \models c_1) & \sqcup_\eta & (\Gamma_2,t \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),t \\
        (\Gamma_1,f \models c_1)  & \sqcup_\eta & (\Gamma_2,f \models c_2)  & = & (\Gamma_1 \sqcup_\eta \Gamma_2),f \\
    \end{array}
\end{equation*}

The $V$ value effectively encodes an additional implicit used element in the context.
We also have some new cases for the semantic function:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        % Added for eta optimization
        (\Gamma_1,f \models c_1) & \coprod_\eta & (\emptyset \models V)    & = & c_1                                                        \\
        (\emptyset \models V)    & \coprod_\eta & (\Gamma_2,f \models c_2) & = & (\emptyset \models CI) \coprod_\eta (\Gamma_2 \models c_2) \\

        (\Gamma_1,t \models c_1) & \coprod_\eta & (\emptyset \models V)    & = & (\Gamma_1 \models (s \coprod_\eta l)) \coprod_\eta i       \\
                                 &              &                          &   & \text{where } l := (\Gamma_1 \models c_1 )                 \\
                                 &              &                          &   & \text{where } s := (\emptyset \models S )                  \\
                                 &              &                          &   & \text{where } i := (\emptyset \models I )                  \\
        (\emptyset \models V)    & \coprod_\eta & (\Gamma_2,t \models c_2) & = & (\emptyset \models SI) \coprod_\eta (\Gamma_2 \models c_2) \\
        (\emptyset \models c_1)  & \coprod_\eta & (\emptyset \models V)    & = & c_1                                                        \\
        (\emptyset \models V)    & \coprod_\eta & (\emptyset \models c_2)  & = & CI c_2                                                     \\
        (\emptyset \models V)    & \coprod_\eta & (\emptyset \models V)    & = & \mathbf{impossible}                                        \\
        \\

        (\Gamma_1 \models c_1)   & \coprod_\eta & (\Gamma_2 \models)       & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models)          \\
    \end{array}
\end{equation*}

The case $(\emptyset \models V) \coprod (\emptyset \models V) $ is impossible in this translation, implementations may throw an error or leave it as undefined behavior.
\textbf{TODO:} Why is this an impossible case?

With the addition of the expression value to $\sqcup_\eta$, the function now takes the same input as $\coprod_\eta$ and has very similar matching patterns.
Our implementation, therefore, folds context merging into $\coprod_\eta$ to make the code more concise.

\subsection{Linear algorithm}
The implementation of the linear algorithm is much simpler because it is based on the $\strict$ algorithm.
The compilation function is identical save for a different semantic function $\coprod_{\text{linear}}$:

\begin{equation*}
    \begin{array}{l l l l}
        \linear & z         & = & 1 \models I                                                              \\
        \linear & s \ e     & = & n+1 \models (0 \models K) \coprod (n \models c)                          \\
                &           &   & \text{where } (n \models c) := \linear e                                 \\
        \\
        \linear & \lambda e & = & 0 \models K \; c                                                         \\
                &           &   & \text{where } (0 \models c) := \linear e                                 \\
        \linear & \lambda e & = & (n-1) \models K \; c                                                     \\
                &           &   & \text{where } (n \models c) := \linear e                                 \\
        \\
        \linear & e_1 \ e_2 & = & (n \models ((n_1 \models c_1) \coprod_{\text{linear}} (n_2 \models c_2)) \\
                &           &   & \text{where } (n_1 \models c_1) := \linear e_1                           \\
                &           &   & \text{where } (n_2 \models c_2) := \linear e_2                           \\
                &           &   & \text{where } n := \max n_1 \, n_2                                       \\
    \end{array}
\end{equation*}

below we give the corresponding semantic function, which is no longer recursive, making this algorithm linear:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        (0 \models c_1)   & \coprod & (0 \models c_2)   & = & c_1 \ c_2                               \\
        \\
        (0 \models c_1)   & \coprod & (n \models c_2)   & = & B_n \ c_1 \ c_2                         \\
        \\
        (n \models c_1)   & \coprod & (0 \models c_2)   & = & C_n \ c_1 \ c_2                         \\
        \\
        (n \models c_1)   & \coprod & (n \models c_2)   & = & S_n \ c_1 \ c_2                         \\
        \\
        (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & B_{n_2-n_1} \ (S_{n_1} \ c_1) \ c_2     \\
                          &         &                   &   & \text{where } n_1 < n_2                 \\
        (n_1 \models c_1) & \coprod & (n_2 \models c_2) & = & C_{n_1-n_2} \ (B_{n_1-n_2} \ c_1) \ c_2 \\
                          &         &                   &   & \text{where } n_1 > n_2                 \\
    \end{array}
\end{equation*}

\textbf{TODO: } Can we add lazy weakening and/or eta optimization?

\section{Miranda}
\label{sec:miranda}
For our base implementation, we try to stick to Miranda's design as much as possible.
While there is published work on Miranda~\cite{turner_miranda_1985}, it does not provide much detail on the reduction engine.
Fortunately, the source code for Miranda is now freely available~\cite{noauthor_open_2021}, and from the code, we can deduce that its design is very similar to SASL~\cite{turner_new_1979}, an earlier language developed by Turner.

Miranda uses a graph reduction engine, so after a program has been converted into combinatory logic using Turner's abstraction algorithm~\cite{turner_another_1979}, it is loaded into a graph structure.
The graph is binary: every node has exactly two children, and the node itself represents the application of the right node to the left one.
Let us start with an example program $SKK \ 3$.
The $S$ combinator takes 3 arguments, but currying lets us write it as $((SK) \; K)\ 3$.
In this form, it is straightforward to construct the associated binary graph, shown in figure~\ref{fig:skk3}.

\begin{figure}
    \includegraphics[width=.23\columnwidth]{skk3}
    \centering
    \caption{$SKK \ 3$ in graph representation.}
    \label{fig:skk3}
\end{figure}

To reduce this graph, we follow the left pointer until we find a combinator to apply.
As we traverse the graph along what is often called the \textit{spine}, we keep the nodes we have seen on a stack, because their right nodes contain the arguments to the combinator we eventually find.

In this example graph, after following two pointers we find the $S$ combinator, and we apply the definition $S f g x \Rightarrow (f x) (g x)$ to rewrite the graph, as shown in figure~\ref{fig:skk3_sred}.

\begin{figure}
    \includegraphics[width=.8\columnwidth]{skk3_sred}
    \centering
    \caption{
        First reduction step for $SKK \ 3$.
        One of the K combinators is underlined to make them easier to distinguish.
    }
    \label{fig:skk3_sred}
\end{figure}

The reduction process continues until the program has been reduced to a constant value, or there are insufficient arguments for the next combinator to reduce.
The expression is said to be in \textit{weak head normal form}: some inner part of the expression may be reducible, but we only perform outermost reduction, so this is as close as we can get to normal form.

\subsection{Sharing}
In the reduction in figure~\ref{fig:skk3_sred}, we have copied the value $3$ to two nodes.
For another reduction where the $x$ value is a pointer to a subgraph, performing the reduction may result in multiple pointers to the same expression.
Expressions in Miranda are \textit{referentially transparent}: we can safely replace any expression with the value it computes.
Computing $x$ in both branches would therefore be wasteful.
Instead, we would like to compute it once (in whatever branch runs first) and have the second branch directly reuse the result.
It turns out that this is not too difficult to implement: when we make a reduction step, we update the outermost node with the contents of the new outermost node.
In figure~\ref{fig:skk3_sred}, this would mean we overwrite node 1 with the contents of node 4.
Now any other pointers to node 1 automatically point to the reduced equivalent.
If the expression reduces to a single value $v$ (say an integer), we can use the $I$ combinator to make an \textit{indirection node} $I v$.

\subsection{Representing the graph}
Miranda represents the graph using three arrays, \texttt{hd} (for head) stores the left pointer, \texttt{tl} (for tail) stores the right pointer, and \texttt{tag} stores metadata for the cell.
Owing to its early inception when 64-bit machines were rare, \texttt{hd} and \texttt{tl} are 32 bits wide.
\texttt{tag} only needs to store a few bit flags and thus is a single byte wide.
Our implementation stores two flags in \texttt{tag}:
\begin{itemize}
    \item \texttt{WANTED}: Set if this cell is currently in use. Cells without this bit set are considered free and may be overwritten to store new nodes in the graph. More on this in the section on Garbage collection.
    \item \texttt{RHS\_INT}: Set if the right pointer is not a pointer but an integer value. Note that the left pointer can never be an integer because the left pointer is always a function for which the right pointer is the first argument.
\end{itemize}

Miranda integers have unlimited precision, using GNU MP~\cite{gnump} internally.
For simplicity (and performance), our implementation instead treats the cell pointer as a 32-bit signed integer.

It is unclear why a representation with three separate arrays was chosen over a single array where each element contains a 'cell' struct with a head, tail and tag.
We suspect decomposing the graph into separate arrays leads to poor data locality, as almost any operation will access all three parts of a cell.
Our base implementation respects this design choice, later implementations do not.

Miranda reserves the lower range of cell pointer values for special purposes:
\begin{itemize}
    \item 0..255 represents the Latin 1 character set.
    \item 256..305 for internal use in the lexer.
    \item 306..446 to encode combinators.
\end{itemize}

The first 446 cells in the heap are unused: when the reduction engine encounters their value, they are interpreted as a character, lexer rule or combinator.
This explains why we have to tag integers but not combinators or character values.

Our implementation does not support character values and needs no reserved number for the lexer, so we can encode the first combinator as $0$.
The bulk combinators introduced by the $\linear$ algorithm do complicate the encoding somewhat:
in principle, there is now an infinite amount of possible combinators.
However, compiled programs only contain a finite subset of those combinators, and the reduction process only creates combinators that were present in the original program.
We can have an efficient encoding of combinators, but only if we generate it for specific programs.
Our implementation keeps an array of function pointers, each pointing to the code to reduce a particular combinator.
The array is populated as the compiled program is loaded into the graph representation:
when we encounter a combinator we have not seen before, its implementation is added to the array, and the combinator is encoded into the graph as its index into the array.
At runtime, combinator resolution is done by indexing into the array and calling the associated function.

\subsection{Strict combinators}
Miranda has special built-in combinators to perform arithmetic operations and compare numbers.
Where most combinators place no restrictions on their arguments, these combinators require their arguments to be fully reduced to integers (hence call them \textit{strict}), and so require special treatment.
We found Miranda's implementation difficult to follow, the approach we describe here is the one taken for our implementation, and may not fully correspond to Miranda's.

When a \textit{strict} combinator is about to be reduced, the engine inspects the arguments.
If all arguments are reduced, the combinator reduction is performed as usual.
If one or more arguments are not yet reduced, we push the pointers to those arguments onto the stack and proceed to reduce from the top of the stack.
Eventually, the top argument will have been fully reduced, producing an integer value.
In general, if a reduction results in an integer value, it can mean one of two things:
\begin{itemize}
    \item The value is the result of the program.
    \item The value is one of the arguments for a strict combinator.
\end{itemize}
We distinguish between these cases by looking at the size of the stack: if it is empty, we have computed the final result of the program.
Otherwise, we continue reducing from the top of the stack, which will be either another argument or the original combinator expression.
The pointer to the original complex argument is replaced with an indirection node to the value, using the update logic described earlier.

As an example, consider the expression $+ \ 1 \ (K \ 2 \ 3)$.
After we have traversed the spine and encountered the $+$ combinator, the engine notices that the second argument is not yet reduced, and adds it to the stack.
Figure~\ref{fig:add_step1} shows the state of the graph at this point in the reduction process.

\begin{figure}
    \includegraphics[width=.8\columnwidth]{add_step1}
    \centering
    \caption{
        Graph of $+ \ 1 \ (K \ 2 \ 3)$ after the second argument has been added to the stack.
    }
    \label{fig:add_step1}
\end{figure}

Once $K \ 2 \ 3$ has been reduced to $3$, the engine observes that the stack is not yet empty, so it creates an indirection node and resumes reduction from the top of the stack.
Figure~\ref{fig:add_step2} reflects this state.

\begin{figure}
    \includegraphics[width=.8\columnwidth]{add_step2}
    \centering
    \caption{
        Graph of $+ \ 1 \ (K \ 2 \ 3)$ after the second argument has been reduced.
    }
    \label{fig:add_step2}
\end{figure}

This example shows that it is not sufficient to check the \texttt{tag} of each argument to see if it is an integer: indirection nodes also count as fully reduced.

\subsection{Garbage collection}
The graph reduction process may lead to some nodes becoming unreachable.
Coming back to the example in figure~\ref{fig:skk3_sred}, after the reduction step nodes 2 and 3 are no longer used: they are \textit{garbage}.
In this example, it is easy to conclude which nodes are garbage, but in larger programs, there are often multiple references to the same node, so we cannot mark nodes as garbage during the reduction step.
Miranda employs a simple mark-and-sweep garbage collector that kicks in when the engine runs out of free cells while allocating a new node.
In the first phase, it zeroes out the \texttt{WANTED} bit on all tags in the heap.
Then in the marking phase, starting from the elements currently on the stack\footnote{
    Miranda keeps its reduction stack on the C stack.
    This makes it difficult to reliably enumerate all cell pointers on the stack at a given point in time, because the compiler may choose to keep some of the pointers in registers.
    For this reason, users are instructed not to enable optimizations when compiling Miranda, which hurts performance but is necessary for correctness.
    When Miranda was developed, compilers were not advanced enough to perform such optimizations, so this was not an issue.
}, the garbage collector traverses the graph and sets the \texttt{WANTED} bit on all reachable nodes.
There is no real sweep phase: instead, the allocation routine walks over the tag array until it finds a cell that is not marked as \texttt{WANTED}.
This also means that there is no list of currently free cells, which we suspect may lead to performance issues in large heaps.

We consider the garbage collector in Miranda to be somewhat simplistic and inefficient, and we suspect there is much to improve in terms of efficiency.
Our benchmark programs do not use too much heap space, so we choose not to focus our efforts on this, and defer optimizations on this front to future work.

\section{TIGRE}
\label{sec:tigre}
The TIGRE engine, first presented in~\cite{koopman_fresh_1989}, is a combinator graph reduction engine like Miranda.
Originally designed for the VAX instruction set, our implementation targets the x86-64 instruction set but mirrors the TIGRE design otherwise.
The TIGRE paper focuses on key differences from Miranda, the elimination of tags and directly executable graph nodes, so where the paper omits implementation details we take the same approach as our reference implementation based on Miranda.

\subsection{Eliminating tags}
The TIGRE authors point to tags as a major source of inefficiency in graph reduction engines.
At evaluation time Miranda must check the tag of nodes to see if their right-hand side is a value or a pointer to a node.
TIGRE does away with this by adding a special \texttt{LIT} combinator.
The \texttt{LIT} combinator takes one argument guaranteed to be a literal value and returns it, like an \texttt{I} combinator specialized for constants.
An apparent downside of this is that simple expressions require more indirection, and therefore a larger graph to reduce.
Take for example the expression $1+2$, which in the Miranda-style engine can be represented with just two nodes, shown in figure~\ref{fig:1_plus_2_miranda}.

\begin{figure}
    \includegraphics[width=.2\columnwidth]{1_plus_2_miranda}
    \centering
    \caption{
        Graph of $1 + 2$ in the Miranda-style engine.
        Tags omitted for brevity.
    }
    \label{fig:1_plus_2_miranda}
\end{figure}

The TIGRE version of this program, shown in figure~\ref{fig:1_plus_2_tigre}, is larger, occupying four nodes.

\begin{figure}
    \includegraphics[width=.46\columnwidth]{1_plus_2_tigre}
    \centering
    \caption{
        Graph of $1 + 2$ in the TIGRE-style engine.
    }
    \label{fig:1_plus_2_tigre}
\end{figure}

In this example, the TIGRE graph may indeed be slower to evaluate, but the same does not hold for the programs in general.
If the values $1$ and $2$ were the results of a previous computation, they would also be behind an indirection node in the Miranda engine.
The TIGRE approach is only slower when performing operations on values that are constant in the source program, in other cases the reduced branching on tag values is expected to speed up execution.

The TIGRE authors report that some engines also use the tag to distinguish between combinators and cell pointers.
In Miranda the lowest cell pointers are reserved for combinators, so this does not need to be encoded in the tag, but the engine still needs to do a case distinction for low values of the pointer.
TIGRE also eliminates this overhead by encoding nodes as executable machine instructions.
We discuss this encoding in more detail in the next section.
With this change any pointer in the graph points to executable code: nodes are directly executable, and for combinators, we can insert a pointer to their implementation function.
The overhead of interpreting tags is thus fully eliminated.

Our implementation still has a tag byte in each cell, but it is exclusively used by the garbage collector to keep track of free and occupied cells.
The TIGRE paper provides no details on garbage collection, so we adopt the same mark-and-sweep collection approach as Miranda.

\subsection{Executable graph nodes}
Due to the elimination of tags, graph nodes must be executable x86-64 machine instructions.
Executing a node is expected to \emph{unwind the spine}: a pointer to the current node should be pushed onto the stack, after which we proceed to evaluate the cell to the left.
It turns out that this can be neatly encoded as a subroutine call in both the VAX and x86-64 instruction sets.

The x86-64 \texttt{CALL} \textit{rel32} instruction\footnote{\url{https://c9x.me/x86/html/file_module_x86_id_26.html}} modifies the processor state as follows:
\begin{enumerate}
    \item Pushes onto the stack a pointer to the instruction directly after the \texttt{CALL}.
          In a regular program, this is the \emph{return address}, the address the program should jump to after the subroutine finishes.
    \item Adds \textit{rel32} to the program counter register \texttt{\%rip}.
          \textit{rel32} encodes the address of the subroutine as a signed 32-bit offset from the location of the \texttt{CALL}.
\end{enumerate}

We exploit this by encoding nodes as a \texttt{CALL} instruction to the left pointer, followed by the right pointer as a raw value.
The memory layout of a node is illustrated in figure~\ref{fig:tigre_layout}.

\begin{figure*}
    \includegraphics[width=2\columnwidth]{tigre_layout}
    \centering
    \caption{
        Memory layout of a node in the TIGRE-style engine.
        Each square represents a byte of memory.
    }
    \label{fig:tigre_layout}
\end{figure*}

When execution jumps to a node, the \texttt{CALL} instruction executes.
The address of the right pointer is pushed to the stack.
This is convenient because when later a combinator executes, it only needs to read the value of the right pointer (the left is only needed to unwind the stack).
Execution then jumps to the address of the left cell pointer, which is either another node or the code for a combinator.
The \texttt{CALL} instruction and 64-bit right pointer take up 13 bytes.
We add an extra byte to store the tag (only one bit is currently used) and 2 bytes of padding so the full node is 16 bytes wide.
Having 16-byte wide nodes is convenient because Intel processors can execute jumps faster if the target is aligned to a 16-byte boundary\footnote{\url{https://groups.google.com/g/golang-nuts/c/dhGbgC1pAmA/m/Rcqwcd5mGmoJ}}.

The simpler combinators are easily expressed as a short sequence of assembly instructions.
Figure~\ref{fig:combI} shows the implementation of the I combinator.

\begin{figure}
    \begin{lstlisting}[
        language={[x86masm]Assembler},
    ]
    comb_I:
        ; Pop pointer to argument
        ; into rax
        pop rax
        ; Dereference and jump
        jmp [rax]
    \end{lstlisting}
    \centering
    \caption{I combinator in x86-64 assembly.}
    \label{fig:combI}
\end{figure}

Combinators that return a value (\texttt{LIT} or strict combinators) can be implemented by computing the value and returning it with \texttt{RET}.
See for example the implementation of \texttt{LIT} in figure~\ref{fig:combLIT}.

\begin{figure}
    \begin{lstlisting}[
        language={[x86masm]Assembler},
    ]
    comb_LIT:
        ; Pop pointer to argument
        ; into rax
        pop rax
        ; Dereference and store 
        ; in return value register
        mov rax, [rax]
        ; Return to caller
        ret
    \end{lstlisting}
    \centering
    \caption{LIT combinator in x86-64 assembly.}
    \label{fig:combLIT}
\end{figure}

A strict combinator that needs to evaluate its arguments is easier to implement in a TIGRE-style engine compared to Miranda: the implementation can strictly evaluate arguments by jumping to the argument node with a \texttt{CALL} instruction, and control will return to the combinator implementation after the argument has been evaluated and the \texttt{RET} instruction executes.
The fully evaluated argument is then available in the \texttt{\%rax} register.
This even works from Rust code: a pointer to a node can be cast to a function pointer with an integer as a return value and invoked normally.

\textbf{TODO: Maybe add a step-by-step reduction example?}

To our surprise, we did not find the TIGRE-style engine more difficult to implement than the Miranda version, mainly because the reduction process happens to map rather naturally to subroutine calls.
We discuss the main challenges encountered during the implementation in the next section.

\subsection{Implementation issues}
\subsubsection{Writeable and executable memory}
For our reduction engine to work, nodes must be stored in executable memory.
Modern x86-64 CPUs support the NX-bit~\cite{paulson_new_2004}, which allows the operating system to mark regions of the memory as non-executable, triggering an exception if a program jumps to a page with the NX-bit set.
The NX-bit is typically enabled by default on dynamically allocated heap memory.
To allocate memory without the NX-bit set on Linux, we allocate memory for the graph using \texttt{mmap}\footnote{\url{https://man7.org/linux/man-pages/man2/mmap.2.html}} with the \texttt{PROT\_EXEC} (to allow execution) and \texttt{PROT\_WRITE} (to allow writing) flags.
Some operating systems do not allow unprivileged programs to set \texttt{PROT\_EXEC} and \texttt{PROT\_WRITE} at the same time.
One solution to this is to flip the memory protection settings from executable to writeable, make the changes, and flip back to executable and continue execution~\cite{de_mooij_wx_2015}.
We anticipate this would be prohibitively expensive for our engine: After a combinator executes, it updates the graph with the new value, so that if the same value is used in a different subgraph, it is only evaluated once (just as Miranda does).
The engine would have to issue two \texttt{mprotect} system calls at every reduction step, a heavy performance penalty.
Luckily, the author's Linux distribution does not enforce a W \^{} X policy and allows the allocation of memory that is both writeable and executable.

\subsubsection{Relative jumps}
The \texttt{CALL} instruction takes as argument a 32-bit offset relative to the instruction pointer (the address of the \texttt{CALL} instruction), rather than a full 64-bit address.
Therefore, some pointer arithmetic is required to calculate the value of this offset.
More importantly, a 32-bit offset is too small to reach any target within the 64-bit address space.
This is problematic for large programs where nodes may be very far apart.
The programs we have used to evaluate the engine are relatively small, so we did not encounter such an issue.
However, the offset may also point to an implementation of a combinator.
The location of this code is determined by the linker when the engine itself is compiled, so it is difficult to influence its location in memory.
To get around this, our engine registers the addresses of all combinators.
Given this set of addresses, the engine can compute the \emph{safe range}: a region of the address space from which a 32-bit offset would be sufficient to reach any combinator.
When allocating memory to store the graph nodes, it finds an unmapped page within the safe range and allocates it.
As an additional safeguard, all offsets are initially computed as 64-bit values, after which the engine verifies that the offset indeed fits in 32 bits.

The technique is not without its problems: for large programs (or a large set of combinators), allocation of large enough block of memory may be unsuccessful.
A more robust solution would be to use trampolines: locations in memory close to nodes that contain assembly instructions to jump to locations far away in memory.
This requires more bookkeeping and is therefore left as future work.

\subsection{Stack alignment}
The System-V ABI that is used on x86-64-Linux mandates that before a \texttt{CALL} instruction the stack is aligned to 16 bytes\footnote{https://c9x.me/compile/doc/abi.html}.
Our engine regularly chains \texttt{CALL} instructions (pushing an 8-byte return address onto the stack) until a combinator is reached, so it is impossible to guarantee statically the alignment of the stack upon a call to a combinator implementation.
The entry point of each combinator that needs to call into Rust code checks the alignment of the stack right before the \texttt{CALL} instruction and adjusts it if necessary.

\subsection{Panicking inside combinators}
Rust has the concept of \emph{panic}: an exceptional situation that indicates a bug in the program, for example, an index out-of-bounds error.
If a panic occurs, the program collects a stack trace and terminates with an error code.
A panic inside of a combinator implementation may cause the program to unwind into the node graph, which is undesirable because it is undefined behaviour\footnote{\url{https://doc.rust-lang.org/nomicon/ffi.html\#ffi-and-unwinding}}.
To prevent this, all combinator implementations are wrapped in a call to \emph{catch\_unwind}\footnote{\url{https://doc.rust-lang.org/std/panic/fn.catch_unwind.html}}, so we catch panics before they leave Rust code, log them and abort the program.

\section{Performance evaluation}
\label{sec:eval}
Programs:
\begin{itemize}
    \item ackermann
    \item digits of e
    \item linfib
    \item lsort
    \item nfib
    \item primes
    \item queens
    \item tak
    \item towers of hanoi
    \item treesort
\end{itemize}

Evaluate performance of:

\begin{itemize}
    \item Bracket abstraction
    \item Kiselyov
    \item GHC
\end{itemize}

Suggestion from \textit{Common Benchmarking Pitfalls}: Always look at your distribution.
When you have run benchmarks, take a look at the times for each sample.

\section{Conclusion}
% Repeat the main question / objective of your research

% Give the answer to the question
Source code for our Superg evaluator is available at \url{https://github.com/wildarch/mono/tree/main/experiments/superg}.

% Why is this the answer? Present the main arguments

% What are the implications of this answer?

% Suggestions for future research

% FINAL CHECK: Can the conclusion be read independently?

\bibliographystyle{IEEEtran}
\bibliography{zotero, references}

\end{document}